Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-06 paper.tex
minor

Thus, a transaction level property follows implication structure -- the 
antecedent of the implication is an event in $\mathcal{ST}$ or 
$\mathcal{HT}$ and the consequent is an event in $\mathcal{HT}$ 
or $\mathcal{ST}$ respectively.  An event in $\mathcal{ST}$ may be 
a function call or an update to program variables. Similarly, an event in 
$\mathcal{HT}$ corresponds to an update of combinational or sequential 
elements. Figure~\ref{figure:transaction} shows an example of transaction 
level property, written as monitors in C, for the UART IP.  In $monitor\_P1()$, 
$outb(UART\_TR,0x0c)$ is a FW event and $ack\_o$ is a HW event. Whereas in
$monitor\_P2()$, the antecedent $tx\_empty()$ is a HW event and $send\_data$ is a FW
event. 
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
\multicolumn{2}{c}{Monitors (in C)}
\\
\hline
\begin{lstlisting}[mathescape=true,language=C,moredelim={[is][keywordstyle]{@@}{@@}}]
int monitor_P1() { 
 if(outb(UART_TR,0x0c)){
  next_timeframe();
  assert(@@ ack_o==1 @@);}}
\end{lstlisting} 
&
\begin{lstlisting}[mathescape=true,language=C,moredelim={[is][keywordstyle]{@@}{@@}}]
int monitor_P2() { 
 assert(!tx_empty() || 
 (@@(send_data&0x1)==1)@@);
}
\end{lstlisting} \\ 
\hline
\end{tabular}
\caption{Transaction-level property as monitor (in C)}
\label{figure:transaction}
\end{figure}
%

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown in the left and the equivalent 
monitor in C is shown in right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
Bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component-level property (in SVA) as monitor (in C)}
\label{figure:component}
\end{figure}
% 
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART IP and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

@@

\textbf{\emph{About UART:}} A UART IP core is used for asynchronous
transmission and reception of data which provides serial communication
capabilities with a modem or other external devices.  The UART IP is
compliant with industry standards for UART and interfaces with the wishbone
bus.  The design statistics for UART is shown in Table~\ref{table:stats}.

\Omit{
The core supports single transmission/reception format which is
1~start bit, 1~stop bit and 8~data bits without parity.  It also contains a
single transmit/receive buffer.  The UART IP implements a baud rate
generation circuit based on a harmonic frequency synthesizer that support a
wide variety of clock frequencies.  The UART IP core is implemented in
approximately 1000 lines of Verilog RTL, generating several thousand gates. 
The UART is obtained from~\url{http://www.opencores.org}.  A software
netlist model (in C) is synthesized from the UART RTL using \emph{v2c}.  The
software netlist model is approximately 1100 LOC.
\rmcmt{refer stats table}
}
%
\begin{table}
\begin{center}
{
\scriptsize
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
  Circuit & Verilog & Latches(L)/ & Input & Output & GATE & Firmware \\
  & LOC & FF & Ports & Ports & Count & (LOC)\\ 
\cline{1-7}
\multicolumn{7}{|c|}{Universal Asynchronous Receiver Transmitter} \\ \hline 
  UART & 1200 & 356 & 12 & 9 & 413 & 528 \\ \hline
\multicolumn{7}{|c|}{System-On-a-Chip} \\ \hline 
  SoC & 3567 & 840 & 14 & 11 & 945 & 734 \\ \hline
\end{tabular}
}
\end{center}
\caption{Design statistics for UART IP and SoC Design}
\label{table:stats}
\end{table}  
%
\textbf{\emph{UART Configuration:}} The UART IP core is configured in
different operating modes, namely--{\em transmission without interrupt
enabled} (Scenario A), {\em transmission with interrupt enabled} (Scenario
B) and {\em loopback mode with interrupt enabled} (Scenario C).  The
data-width varies in each mode, ranging from 8~bits to 64~bits.  Scenario A
and Scenario B exercises only the transmitter module to transmit
non-deterministic data through the serial output.  The receiver module is
inactive in Scenario A and Scenario B.  By contrast, in Scenario C we
configure the UART in loopback mode with interrupt logic enabled.  Thus,
both the transmitter and the receiver are active in this mode.  The firmware
can receive interrupt from the transmitter or receiver module depending on
whether the transmitter buffer is empty or the receiver buffer is full
respectively, hence the control is non-deterministic in this mode. \\
%
\textbf{\emph{Discussion:}}
Table~\ref{table:safe} reports the run times for bounded safety proofs of
FW-HW interaction properties of the UART IP core.  Column~1 in
Table~\ref{table:safe} gives the name of the scenario -- transmission
without interrupt (transmit), transmission with interrupt (trans\_intr) and
loopback mode (loopback).  Column~2 gives the maximum loop bound in the
firmware.  Column~3~and~4 present the co-verification times using \hwcbmcv
at the bit-level and word-level.  Column~7 gives the verification
times using \verifox.  Column~5~and~6 report
the total/feasible path counts and the percentage of infeasible paths pruned
for each mode.  The focus of our experiment is to compare the performance of 
path-based techniques over monolithic BMC. 
%effectively pruning infeasible FW-HW interactions with respect to a
%scenario.  is demonstrated by the \%-age of pruned paths (reported in
%column~9) for each scenario.

\textbf{\emph{Safety properties:}}
We verify several properties of the FW-HW interactions of UART IP. 
In Scenario~A and scenario~B, we verified whether
the transmitted data (32-bit or 64-bit) is available through the serial
output port after a pre-determined number of clock cycles.
In both of the
configurations, \verifox is able to prune the receiver logic since the
FW only exercises the transmitter module by appropriately configuring
the memory-mapped registers.  In Scenario~C, we verified whether the transmitted
data matches the data received when the UART is configured in loopback
mode.\\ 
\textbf{\emph{Detecting UART bugs:}}
We found several bugs in the open source UART IP obtained from~\url{http://www.opencores.org}.  
The bottom part of Table~\ref{table:safe} reports the verification
times for detecting data-path and control-path bugs in UART IP.  
The bug in the UART data-path is manifested in the transmission mode when no
interrupt is enabled.  The bug occurs when the transmitted data overlaps
with the start and stop bits resulting in incorrect bit pattern in the
serial output of the transmitter.  However, when the interrupt is enabled
during transmission mode, the bug in the control-path of the transmitter
module is triggered which deactivates the transmitter $empty$ signal.  This
illegally prevents any transmitter interrupt to occur even when the
transmitter buffer is empty, thereby violating the data transfer protocol. 
On the other hand, the \emph{control bug} occurs in the loopback mode only when the
data is present at the receiver buffer but the $data\_present$ signal is
never asserted, as a result of which the receiver interrupt is never issued
in the FW. Thus, the transmitted data is not the same as the received
data in loopback mode leading to a control bug.

@@

\begin{table}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{2}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} \\ 
\cline{3-7}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & \multicolumn{3}{c|}{VerifOx} \\ 
\cline{3-7}
      &       &  Bit-level & Word-level & Total/Feasible & \%-age & $\mathcal{M}_{seq}$  \\
%\cline{1-7}
      &       &   Time     &   Time      & Paths & Pruning & Time \\
\cline{1-7}      
\multicolumn{7}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
  transmit (32) & 250 & 15.02 & 15.78 & 247104/224 & 99.90 & \textbf{1}.{13} \\ 
transmit (64) & 500 & 23.87 & 23.29 & 247104/324 & 99.86 & \textbf{1}.\textbf{61}  \\ \hline 
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
  trans\_intr (32) & 250 & 14.86 & 15.98 & 247104/295 & 99.88 & \textbf{1}.{49} \\
  trans\_intr (64) & 500 & 24.14 & 26.42 & 247104/362 & 99.85 & \textbf{1}.{81} \\ \hline
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
loopback (8)  & 230 & 52.06 & 55.71 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} \\ 
loopback (16) & 500 & 122.12 & 147.65 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} \\ 
loopback (32) & 650 & 170.62 & 192.35 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85}   \\ 
loopback (64) & 1300 & 409.71 & 445.66 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}  \\ \hline
\multicolumn{7}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
  transmit (64) & 520 & 28.43 & 28.14 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{12} \\ \hline
\multicolumn{7}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 29.70 & 247104/362 & 99.85 & \textbf{1}.\textbf{05} \\ \hline
\multicolumn{7}{|c|}{detecting control bugs in loopback mode} \\ \hline
loopback (64) & 1300 & 443.15 & 427.68 & 247104/2566 & 98.96 & \textbf{62}.\textbf{34} \\ \hline
\end{tabular}
\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of UART 
\label{table:safe}}
\end{table}

@@

\begin{table}
\begin{center}
{
%\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{2}{c|}{Monolithic} & \multicolumn{1}{c|}{Path-based} \\ 
\cline{3-5}
 Scenario &  & \multicolumn{2}{c|}{\hwcbmcv} & \verifox \\ 
\cline{3-5}
      &       &  Bit-level & Word-level & $\mathcal{M}_{seq}$ \\
\cline{1-5}
      &       &   Time     &   Time     & Time \\
\cline{1-5}      
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control (safe)} \\ \hline
data\_transfer & 20 & 86.92 & 86.23 & \textbf{17}.\textbf{42} \\ 
\hline
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control
(unsafe)} \\ \hline
data\_transfer & 20 & 92.63 & 91.28 & \textbf{1}.\textbf{27} \\ 
\hline
\end{tabular}
%\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of a SoC 
\label{table:SoC}}
\end{table}

>>>>

Thus, a transaction level property follows implication structure -- the 
antecedent of the implication is an event in $\mathcal{ST}$ or 
$\mathcal{HT}$ and the consequent is an event in $\mathcal{HT}$ 
or $\mathcal{ST}$ respectively.  An event in $\mathcal{ST}$ may be 
a function call or an update to program variables. Similarly, an event in 
$\mathcal{HT}$ corresponds to an update of combinational or sequential 
elements. Figure~\ref{figure:transaction} shows an example of transaction 
level property, written as monitors in C, for the UART IP.  In $monitor\_P1()$, 
$outb(UART\_TR,0x0c)$ is a FW event and $ack\_o$ is a HW event. Whereas in
$monitor\_P2()$, the antecedent $tx\_empty()$ is a HW event and $send\_data$ is a FW
event. 
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
\multicolumn{2}{c}{Monitors (in C)}
\\
\hline
\begin{lstlisting}[mathescape=true,language=C,moredelim={[is][keywordstyle]{@@}{@@}}]
bool monitor_P1() { 
 if(outb(UART_TR,0x0c)){
  next_timeframe();
  assert(@@ ack_o==1 @@);}}
\end{lstlisting} 
&
\begin{lstlisting}[mathescape=true,language=C,moredelim={[is][keywordstyle]{@@}{@@}}]
bool monitor_P2() { 
 assert(!tx_empty() || 
 (@@(send_data&0x1)==1)@@);
}
\end{lstlisting} \\ 
\hline
\end{tabular}
\caption{Transaction level property as monitor (in C)}
\label{figure:transaction}
\end{figure}
%

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown in the left and the equivalent 
monitor in C is shown in right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component level property as monitor (in C)}
\label{figure:component}
\end{figure}
% 
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART IP and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

@@

\textbf{\emph{About UART:}} A UART IP core is used for asynchronous
transmission and reception of data which provides serial communication
capabilities with a modem or other external devices.  The UART IP is
compliant with industry standards for UART and interfaces with the wishbone
bus.  The design statistics for UART is shown in Table~\ref{table:stats}.
\begin{table}
\begin{center}
{
\begin{tabular}{|c|c|c|c|c|c|c|}
  Circuit & Verilog & Latches(L)/ & Input & Output & GATE & Firmware \\
  & LOC & FF & Ports & Ports & Count & (LOC)\\ 
\cline{1-7}
\multicolumn{7}{|c|}{Universal Asynchronous Receiver Transmitter} \\ \hline 
  UART & 1200 & 356 & 12 & 9 & 413 & 528 \\ \hline
\multicolumn{7}{|c|}{System-On-a-Chip} \\ \hline 
  SoC & 3567 & 840 & 14 & 11 & 945 & 734 \\ \hline
\end{center}
\caption{Design statistics for UART IP and SoC Design}
\label{table:stats}
\end{table}  
\textbf{\emph{UART Configuration:}} The UART IP core is configured in
different operating modes, namely--{\em transmission without interrupt
enabled} (Scenario A), {\em transmission with interrupt enabled} (Scenario
B) and {\em loopback mode with interrupt enabled} (Scenario C).  The
data-width varies in each mode, ranging from 8~bits to 64~bits.  Scenario A
and Scenario B exercises only the transmitter module to transmit
non-deterministic data through the serial output.  The receiver module is
inactive in Scenario A and Scenario B.  By contrast, in Scenario C we
configure the UART in loopback mode with interrupt logic enabled.  Thus,
both the transmitter and the receiver are active in this mode.  The firmware
can receive interrupt from the transmitter or receiver module depending on
whether the transmitter buffer is empty or the receiver buffer is full
respectively, hence the control is non-deterministic in this mode. \\
\textbf{\emph{Discussion:}}
Table~\ref{table:safe} reports the run times for bounded safety proofs of
FW-HW interaction properties of the UART IP core.  Column~1 in
Table~\ref{table:safe} gives the name of the scenario -- transmission
without interrupt (transmit), transmission with interrupt (trans\_intr) and
loopback mode (loopback).  Column~2 gives the maximum loop bound in the
firmware.  Column~3~and~4 present the co-verification times using \hwcbmcv
at the bit-level and word-level.  Column~7 gives the verification
times using \verifox.  Column~5~and~6 report
the total/feasible path counts and the percentage of infeasible paths pruned
for each mode.  The focus of our experiment is to compare the performance of 
path-based techniques over monolithic BMC.\\ 
\textbf{\emph{Safety properties:}}
We verify several properties of the FW-HW interactions of UART IP. 
In Scenario~A and scenario~B, we verified whether
the transmitted data (32-bit or 64-bit) is available through the serial
output port after a pre-determined number of clock cycles.
In both of the
configurations, \verifox is able to prune the receiver logic since the
FW only exercises the transmitter module by appropriately configuring
the memory-mapped registers.  In Scenario~C, we verified whether the transmitted
data matches the data received when the UART is configured in loopback
mode.\\ 
\textbf{\emph{Detecting UART bugs:}}
We found several bugs in the open source UART IP obtained from~\url{http://www.opencores.org}.  
The bottom part of Table~\ref{table:safe} reports the verification
times for detecting data-path and control-path bugs in UART IP.  
The bug in the UART data-path is manifested in the transmission mode when no
interrupt is enabled.  The bug occurs when the transmitted data overlaps
with the start and stop bits resulting in incorrect bit pattern in the
serial output of the transmitter.  However, when the interrupt is enabled
during transmission mode, the bug in the control-path of the transmitter
module is triggered which deactivates the transmitter $empty$ signal.  This
illegally prevents any transmitter interrupt to occur even when the
transmitter buffer is empty, thereby violating the data transfer protocol. 
On the other hand, the \emph{control bug} occurs in the loopback mode only when the
data is present at the receiver buffer but the $data\_present$ signal is
never asserted, as a result of which the receiver interrupt is never issued
in the FW. Thus, the transmitted data is not the same as the received
data in loopback mode leading to a control bug.

@@

\begin{table}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{2}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} \\ 
\cline{3-7}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & \multicolumn{3}{c|}{\verifox} \\ 
\cline{3-7}
      &       &  Bit-level & Word-level & Total/Feasible & \%-age & $\mathcal{M}_{seq}$  \\
%\cline{1-7}
      &       &   Time     &   Time      & Paths & Pruning & Time \\
\cline{1-7}      
\multicolumn{7}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
  transmit (32) & 250 & 15.02 & 15.78 & 247104/224 & 99.90 & \textbf{1}.{13} \\ 
transmit (64) & 500 & 23.87 & 23.29 & 247104/324 & 99.86 & \textbf{1}.\textbf{61}  \\ \hline 
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
  trans\_intr (32) & 250 & 14.86 & 15.98 & 247104/295 & 99.88 & \textbf{1}.{49} \\
  trans\_intr (64) & 500 & 24.14 & 26.42 & 247104/362 & 99.85 & \textbf{1}.{81} \\ \hline
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
loopback (8)  & 230 & 52.06 & 55.71 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} \\ 
loopback (16) & 500 & 122.12 & 147.65 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} \\ 
loopback (32) & 650 & 170.62 & 192.35 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85}   \\ 
loopback (64) & 1300 & 409.71 & 445.66 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}  \\ \hline
\multicolumn{7}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
  transmit (64) & 520 & 28.43 & 28.14 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{12} \\ \hline
\multicolumn{7}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 29.70 & 247104/362 & 99.85 & \textbf{1}.\textbf{05} \\ \hline
\multicolumn{7}{|c|}{detecting control bugs in loopback mode} \\ \hline
loopback (64) & 1300 & 443.15 & 427.68 & 247104/2566 & 98.96 & \textbf{62}.\textbf{34} \\ \hline
\end{tabular}
\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for verification of UART 
\label{table:safe}}
\end{table}

@@

\begin{table}
\begin{center}
{
%\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{2}{c|}{Monolithic} & \multicolumn{1}{c|}{Path-based} \\ 
\cline{3-5}
 Scenario &  & \multicolumn{2}{c|}{\hwcbmcv} & \verifox \\ 
\cline{3-5}
      &       &  Bit-level & Word-level & $\mathcal{M}_{seq}$ \\
\cline{1-5}
      &       &   Time     &   Time     & Time \\
\cline{1-5}      
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control (safe)} \\ \hline
data\_transfer & 20 & 86.92 & 86.23 & \textbf{17}.\textbf{42} \\ 
\hline
\multicolumn{5}{|c|}{non-deterministic data and non-deterministic control
(unsafe)} \\ \hline
data\_transfer & 20 & 92.63 & 91.28 & \textbf{1}.\textbf{27} \\ 
\hline
\end{tabular}
%\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Bounded safety proof times for verification of a SoC 
\label{table:SoC}}
\end{table}
